---
title: "DATA 607 - Assignment #2"
author: "CLAUDIO, Mauricio"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
---

## Importing a CSV file into R

SQL database files can be exported as Excel files, and then imported into R as comma-separated value (CSV) files using the `read.csv` command. 

```{r fig.width=6, fig.height=6, fig.align='center'}
report.csv=read.csv("C:\\Users\\LENOVO\\OneDrive\\Learning\\Courses In Progress\\DATA 607\\Week 2\\Assignment 2\\report.csv",header=TRUE)
report.csv
```




## Importing an SQL database into R

SQL databases or individual tables can also be imported into R using the RODBC package. The first step is to install the package and load in the library. Then a connection is made to SQL Server by creating an ODBC connection named here "MovieReviews". Queries of the database in the "MovieReviews" connection are made directly in R with the `sqlQuery` command.

```{r fig.width=6, fig.height=6, fig.align='center'}
library(RODBC)
connection=odbcConnect("MovieReviews")
report.sql=sqlQuery(connection,
          "SELECT *
          FROM report
          ORDER BY [Movie Title],Rating DESC")
report.sql
```

## Code

The SQL code for the 'MovieReviews' database and 'report' table is found [here](https://github.com/MauricioClaudio/CUNY-MPS-DATA-607/blob/b868dc70b7b5a6411f7864d844b75669f1811083/Assignment%202%20-%20Mauricio%20Claudio.sql)
