---
title: "DATA 607 - Project 1"
author: "CLAUDIO, Mauricio"
date: "Due date: 2021-09-19"
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: vignette
---


In this project, you’re given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) 
with the following information for all of the players:  
  
Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents  
  
For the first player, the information would be:
Gary Hua, ON, 6.0, 1794, 1605  
  
1605 was calculated by using the pre-tournament opponents’ ratings of 1436, 1563, 1600, 1610, 1649, 1663, 1716, and dividing by the total number of games played.

### Load R libraries and source data

```{r}
library(stringr)
tour.info=
'-----------------------------------------------------------------------------------------
 Pair | Player Name                     |Total|Round|Round|Round|Round|Round|Round|Round| 
 Num  | USCF ID / Rtg (Pre->Post)       | Pts |  1  |  2  |  3  |  4  |  5  |  6  |  7  | 
-----------------------------------------------------------------------------------------
    1 | GARY HUA                        |6.0  |W  39|W  21|W  18|W  14|W   7|D  12|D   4|
   ON | 15445895 / R: 1794   ->1817     |N:2  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
    2 | DAKSHESH DARURI                 |6.0  |W  63|W  58|L   4|W  17|W  16|W  20|W   7|
   MI | 14598900 / R: 1553   ->1663     |N:2  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
    3 | ADITYA BAJAJ                    |6.0  |L   8|W  61|W  25|W  21|W  11|W  13|W  12|
   MI | 14959604 / R: 1384   ->1640     |N:2  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
    4 | PATRICK H SCHILLING             |5.5  |W  23|D  28|W   2|W  26|D   5|W  19|D   1|
   MI | 12616049 / R: 1716   ->1744     |N:2  |W    |B    |W    |B    |W    |B    |B    |
-----------------------------------------------------------------------------------------
    5 | HANSHI ZUO                      |5.5  |W  45|W  37|D  12|D  13|D   4|W  14|W  17|
   MI | 14601533 / R: 1655   ->1690     |N:2  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
    6 | HANSEN SONG                     |5.0  |W  34|D  29|L  11|W  35|D  10|W  27|W  21|
   OH | 15055204 / R: 1686   ->1687     |N:3  |W    |B    |W    |B    |B    |W    |B    |
-----------------------------------------------------------------------------------------
    7 | GARY DEE SWATHELL               |5.0  |W  57|W  46|W  13|W  11|L   1|W   9|L   2|
   MI | 11146376 / R: 1649   ->1673     |N:3  |W    |B    |W    |B    |B    |W    |W    |
-----------------------------------------------------------------------------------------
    8 | EZEKIEL HOUGHTON                |5.0  |W   3|W  32|L  14|L   9|W  47|W  28|W  19|
   MI | 15142253 / R: 1641P17->1657P24  |N:3  |B    |W    |B    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
    9 | STEFANO LEE                     |5.0  |W  25|L  18|W  59|W   8|W  26|L   7|W  20|
   ON | 14954524 / R: 1411   ->1564     |N:2  |W    |B    |W    |B    |W    |B    |B    |
-----------------------------------------------------------------------------------------
   10 | ANVIT RAO                       |5.0  |D  16|L  19|W  55|W  31|D   6|W  25|W  18|
   MI | 14150362 / R: 1365   ->1544     |N:3  |W    |W    |B    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   11 | CAMERON WILLIAM MC LEMAN        |4.5  |D  38|W  56|W   6|L   7|L   3|W  34|W  26|
   MI | 12581589 / R: 1712   ->1696     |N:3  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   12 | KENNETH J TACK                  |4.5  |W  42|W  33|D   5|W  38|H    |D   1|L   3|
   MI | 12681257 / R: 1663   ->1670     |N:3  |W    |B    |W    |B    |     |W    |B    |
-----------------------------------------------------------------------------------------
   13 | TORRANCE HENRY JR               |4.5  |W  36|W  27|L   7|D   5|W  33|L   3|W  32|
   MI | 15082995 / R: 1666   ->1662     |N:3  |B    |W    |B    |B    |W    |W    |B    |
-----------------------------------------------------------------------------------------
   14 | BRADLEY SHAW                    |4.5  |W  54|W  44|W   8|L   1|D  27|L   5|W  31|
   MI | 10131499 / R: 1610   ->1618     |N:3  |W    |B    |W    |W    |B    |B    |W    |
-----------------------------------------------------------------------------------------
   15 | ZACHARY JAMES HOUGHTON          |4.5  |D  19|L  16|W  30|L  22|W  54|W  33|W  38|
   MI | 15619130 / R: 1220P13->1416P20  |N:3  |B    |B    |W    |W    |B    |B    |W    |
-----------------------------------------------------------------------------------------
   16 | MIKE NIKITIN                    |4.0  |D  10|W  15|H    |W  39|L   2|W  36|U    |
   MI | 10295068 / R: 1604   ->1613     |N:3  |B    |W    |     |B    |W    |B    |     |
-----------------------------------------------------------------------------------------
   17 | RONALD GRZEGORCZYK              |4.0  |W  48|W  41|L  26|L   2|W  23|W  22|L   5|
   MI | 10297702 / R: 1629   ->1610     |N:3  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   18 | DAVID SUNDEEN                   |4.0  |W  47|W   9|L   1|W  32|L  19|W  38|L  10|
   MI | 11342094 / R: 1600   ->1600     |N:3  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   19 | DIPANKAR ROY                    |4.0  |D  15|W  10|W  52|D  28|W  18|L   4|L   8|
   MI | 14862333 / R: 1564   ->1570     |N:3  |W    |B    |W    |B    |W    |W    |B    |
-----------------------------------------------------------------------------------------
   20 | JASON ZHENG                     |4.0  |L  40|W  49|W  23|W  41|W  28|L   2|L   9|
   MI | 14529060 / R: 1595   ->1569     |N:4  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   21 | DINH DANG BUI                   |4.0  |W  43|L   1|W  47|L   3|W  40|W  39|L   6|
   ON | 15495066 / R: 1563P22->1562     |N:3  |B    |W    |B    |W    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   22 | EUGENE L MCCLURE                |4.0  |W  64|D  52|L  28|W  15|H    |L  17|W  40|
   MI | 12405534 / R: 1555   ->1529     |N:4  |W    |B    |W    |B    |     |W    |B    |
-----------------------------------------------------------------------------------------
   23 | ALAN BUI                        |4.0  |L   4|W  43|L  20|W  58|L  17|W  37|W  46|
   ON | 15030142 / R: 1363   ->1371     |     |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   24 | MICHAEL R ALDRICH               |4.0  |L  28|L  47|W  43|L  25|W  60|W  44|W  39|
   MI | 13469010 / R: 1229   ->1300     |N:4  |B    |W    |B    |B    |W    |W    |B    |
-----------------------------------------------------------------------------------------
   25 | LOREN SCHWIEBERT                |3.5  |L   9|W  53|L   3|W  24|D  34|L  10|W  47|
   MI | 12486656 / R: 1745   ->1681     |N:4  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   26 | MAX ZHU                         |3.5  |W  49|W  40|W  17|L   4|L   9|D  32|L  11|
   ON | 15131520 / R: 1579   ->1564     |N:4  |B    |W    |B    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
   27 | GAURAV GIDWANI                  |3.5  |W  51|L  13|W  46|W  37|D  14|L   6|U    |
   MI | 14476567 / R: 1552   ->1539     |N:4  |W    |B    |W    |B    |W    |B    |     |
-----------------------------------------------------------------------------------------
   28 | SOFIA ADINA STANESCU-BELLU      |3.5  |W  24|D   4|W  22|D  19|L  20|L   8|D  36|
   MI | 14882954 / R: 1507   ->1513     |N:3  |W    |W    |B    |W    |B    |B    |W    |
-----------------------------------------------------------------------------------------
   29 | CHIEDOZIE OKORIE                |3.5  |W  50|D   6|L  38|L  34|W  52|W  48|U    |
   MI | 15323285 / R: 1602P6 ->1508P12  |N:4  |B    |W    |B    |W    |W    |B    |     |
-----------------------------------------------------------------------------------------
   30 | GEORGE AVERY JONES              |3.5  |L  52|D  64|L  15|W  55|L  31|W  61|W  50|
   ON | 12577178 / R: 1522   ->1444     |     |W    |B    |B    |W    |W    |B    |B    |
-----------------------------------------------------------------------------------------
   31 | RISHI SHETTY                    |3.5  |L  58|D  55|W  64|L  10|W  30|W  50|L  14|
   MI | 15131618 / R: 1494   ->1444     |     |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   32 | JOSHUA PHILIP MATHEWS           |3.5  |W  61|L   8|W  44|L  18|W  51|D  26|L  13|
   ON | 14073750 / R: 1441   ->1433     |N:4  |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   33 | JADE GE                         |3.5  |W  60|L  12|W  50|D  36|L  13|L  15|W  51|
   MI | 14691842 / R: 1449   ->1421     |     |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   34 | MICHAEL JEFFERY THOMAS          |3.5  |L   6|W  60|L  37|W  29|D  25|L  11|W  52|
   MI | 15051807 / R: 1399   ->1400     |     |B    |W    |B    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   35 | JOSHUA DAVID LEE                |3.5  |L  46|L  38|W  56|L   6|W  57|D  52|W  48|
   MI | 14601397 / R: 1438   ->1392     |     |W    |W    |B    |W    |B    |B    |W    |
-----------------------------------------------------------------------------------------
   36 | SIDDHARTH JHA                   |3.5  |L  13|W  57|W  51|D  33|H    |L  16|D  28|
   MI | 14773163 / R: 1355   ->1367     |N:4  |W    |B    |W    |B    |     |W    |B    |
-----------------------------------------------------------------------------------------
   37 | AMIYATOSH PWNANANDAM            |3.5  |B    |L   5|W  34|L  27|H    |L  23|W  61|
   MI | 15489571 / R:  980P12->1077P17  |     |     |B    |W    |W    |     |B    |W    |
-----------------------------------------------------------------------------------------
   38 | BRIAN LIU                       |3.0  |D  11|W  35|W  29|L  12|H    |L  18|L  15|
   MI | 15108523 / R: 1423   ->1439     |N:4  |W    |B    |W    |W    |     |B    |B    |
-----------------------------------------------------------------------------------------
   39 | JOEL R HENDON                   |3.0  |L   1|W  54|W  40|L  16|W  44|L  21|L  24|
   MI | 12923035 / R: 1436P23->1413     |N:4  |B    |W    |B    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
   40 | FOREST ZHANG                    |3.0  |W  20|L  26|L  39|W  59|L  21|W  56|L  22|
   MI | 14892710 / R: 1348   ->1346     |     |B    |B    |W    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
   41 | KYLE WILLIAM MURPHY             |3.0  |W  59|L  17|W  58|L  20|X    |U    |U    |
   MI | 15761443 / R: 1403P5 ->1341P9   |     |B    |W    |B    |W    |     |     |     |
-----------------------------------------------------------------------------------------
   42 | JARED GE                        |3.0  |L  12|L  50|L  57|D  60|D  61|W  64|W  56|
   MI | 14462326 / R: 1332   ->1256     |     |B    |W    |B    |B    |W    |W    |B    |
-----------------------------------------------------------------------------------------
   43 | ROBERT GLEN VASEY               |3.0  |L  21|L  23|L  24|W  63|W  59|L  46|W  55|
   MI | 14101068 / R: 1283   ->1244     |     |W    |B    |W    |W    |B    |B    |W    |
-----------------------------------------------------------------------------------------
   44 | JUSTIN D SCHILLING              |3.0  |B    |L  14|L  32|W  53|L  39|L  24|W  59|
   MI | 15323504 / R: 1199   ->1199     |     |     |W    |B    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   45 | DEREK YAN                       |3.0  |L   5|L  51|D  60|L  56|W  63|D  55|W  58|
   MI | 15372807 / R: 1242   ->1191     |     |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   46 | JACOB ALEXANDER LAVALLEY        |3.0  |W  35|L   7|L  27|L  50|W  64|W  43|L  23|
   MI | 15490981 / R:  377P3 ->1076P10  |     |B    |W    |B    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
   47 | ERIC WRIGHT                     |2.5  |L  18|W  24|L  21|W  61|L   8|D  51|L  25|
   MI | 12533115 / R: 1362   ->1341     |     |W    |B    |W    |B    |W    |B    |W    |
-----------------------------------------------------------------------------------------
   48 | DANIEL KHAIN                    |2.5  |L  17|W  63|H    |D  52|H    |L  29|L  35|
   MI | 14369165 / R: 1382   ->1335     |     |B    |W    |     |B    |     |W    |B    |
-----------------------------------------------------------------------------------------
   49 | MICHAEL J MARTIN                |2.5  |L  26|L  20|D  63|D  64|W  58|H    |U    |
   MI | 12531685 / R: 1291P12->1259P17  |     |W    |W    |B    |W    |B    |     |     |
-----------------------------------------------------------------------------------------
   50 | SHIVAM JHA                      |2.5  |L  29|W  42|L  33|W  46|H    |L  31|L  30|
   MI | 14773178 / R: 1056   ->1111     |     |W    |B    |W    |B    |     |B    |W    |
-----------------------------------------------------------------------------------------
   51 | TEJAS AYYAGARI                  |2.5  |L  27|W  45|L  36|W  57|L  32|D  47|L  33|
   MI | 15205474 / R: 1011   ->1097     |     |B    |W    |B    |W    |B    |W    |W    |
-----------------------------------------------------------------------------------------
   52 | ETHAN GUO                       |2.5  |W  30|D  22|L  19|D  48|L  29|D  35|L  34|
   MI | 14918803 / R:  935   ->1092     |N:4  |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   53 | JOSE C YBARRA                   |2.0  |H    |L  25|H    |L  44|U    |W  57|U    |
   MI | 12578849 / R: 1393   ->1359     |     |     |B    |     |W    |     |W    |     |
-----------------------------------------------------------------------------------------
   54 | LARRY HODGE                     |2.0  |L  14|L  39|L  61|B    |L  15|L  59|W  64|
   MI | 12836773 / R: 1270   ->1200     |     |B    |B    |W    |     |W    |B    |W    |
-----------------------------------------------------------------------------------------
   55 | ALEX KONG                       |2.0  |L  62|D  31|L  10|L  30|B    |D  45|L  43|
   MI | 15412571 / R: 1186   ->1163     |     |W    |B    |W    |B    |     |W    |B    |
-----------------------------------------------------------------------------------------
   56 | MARISA RICCI                    |2.0  |H    |L  11|L  35|W  45|H    |L  40|L  42|
   MI | 14679887 / R: 1153   ->1140     |     |     |B    |W    |W    |     |B    |W    |
-----------------------------------------------------------------------------------------
   57 | MICHAEL LU                      |2.0  |L   7|L  36|W  42|L  51|L  35|L  53|B    |
   MI | 15113330 / R: 1092   ->1079     |     |B    |W    |W    |B    |W    |B    |     |
-----------------------------------------------------------------------------------------
   58 | VIRAJ MOHILE                    |2.0  |W  31|L   2|L  41|L  23|L  49|B    |L  45|
   MI | 14700365 / R:  917   -> 941     |     |W    |B    |W    |B    |W    |     |B    |
-----------------------------------------------------------------------------------------
   59 | SEAN M MC CORMICK               |2.0  |L  41|B    |L   9|L  40|L  43|W  54|L  44|
   MI | 12841036 / R:  853   -> 878     |     |W    |     |B    |B    |W    |W    |B    |
-----------------------------------------------------------------------------------------
   60 | JULIA SHEN                      |1.5  |L  33|L  34|D  45|D  42|L  24|H    |U    |
   MI | 14579262 / R:  967   -> 984     |     |W    |B    |B    |W    |B    |     |     |
-----------------------------------------------------------------------------------------
   61 | JEZZEL FARKAS                   |1.5  |L  32|L   3|W  54|L  47|D  42|L  30|L  37|
   ON | 15771592 / R:  955P11-> 979P18  |     |B    |W    |B    |W    |B    |W    |B    |
-----------------------------------------------------------------------------------------
   62 | ASHWIN BALAJI                   |1.0  |W  55|U    |U    |U    |U    |U    |U    |
   MI | 15219542 / R: 1530   ->1535     |     |B    |     |     |     |     |     |     |
-----------------------------------------------------------------------------------------
   63 | THOMAS JOSEPH HOSMER            |1.0  |L   2|L  48|D  49|L  43|L  45|H    |U    |
   MI | 15057092 / R: 1175   ->1125     |     |W    |B    |W    |B    |B    |     |     |
-----------------------------------------------------------------------------------------
   64 | BEN LI                          |1.0  |L  22|D  30|L  31|D  49|L  46|L  42|L  54|
   MI | 15006561 / R: 1163   ->1112     |     |B    |W    |W    |B    |W    |B    |B    |
-----------------------------------------------------------------------------------------
'
```



### Extract and tidy up player names

```{r}
patrn.NAME="([A-Z]+.){3}[A-Z]*"
NAME=unlist(str_extract_all(tour.info,patrn.NAME))
NAME=NAME[-1]
NAME=trimws(NAME,c("right"))
NAME
```
 
### Extract and tidy up player states

```{r}
patrn.STATE="(\\s){3}[A-Z]+."
STATE=unlist(str_extract_all(tour.info,patrn.STATE))
STATE=STATE[-1]
STATE=trimws(trimws(STATE,c("left")),c("right"))
STATE
```
  
  
### Extract and tidy up player points  

```{r}
patrn.POINTS="\\|[0-9]+.."
POINTS=unlist(str_extract_all(tour.info,patrn.POINTS))
POINTS=str_remove(POINTS,"\\|")
POINTS
```
  
  
### Extract and tidy up player pre-ratings    

```{r}
patrn.PRERATING="R: [0-9]*...."
PRERATING=unlist(str_extract_all(tour.info,patrn.PRERATING))
PRERATING=str_remove_all(PRERATING,"R:")
PRERATING=str_remove_all(PRERATING,"-")
PRERATING=trimws(trimws(PRERATING,c("left")),c("right"))
PRERATING=str_remove_all(PRERATING,"[A-Z][0-9]+")
PRERATING=as.integer(PRERATING)
PRERATING
```
  
  
### Extract and tidy up player opponent index number
  
 
```{r}
patrn.OPPS="(\\d{1,2}|\\s{2})\\|"
OPPS=unlist(str_extract_all(tour.info,patrn.OPPS))
OPPS=OPPS[-(1:10)]
OPPS=str_replace_all(OPPS,"\\|","")
OPPS=as.integer(OPPS)
OPPS
```
  
The resulting vector contains repeating patterns of two empty elements, seven elements of opponent indices and nine empty elements. The seven indices of interest are separated from the two empty elements to the left and the nine empty elements to the right by casting the vector onto a matrix of 18 columns. Then, empty columns 1-2 and 10-18 are eliminated by subsetting columns 3-9, leaving behind only the seven opponent indices of interest. 
  
```{r}
OPPS=matrix(OPPS,ncol=18,byrow=TRUE)
OPPS=OPPS[,3:9]
OPPS
```
  

### Create a data-frame with the extracted variables
  
```{r}
reslt=data.frame(NAME,STATE,POINTS,PRERATING,OPPS,stringsAsFactors=FALSE)
colnames(reslt)[5:11]=c("Opp1","Opp2","Opp3","Opp4","Opp5","Opp6","Opp7")
reslt
```
  
  
### Calculate and store average opponent rating
The ```mean()``` function does not play nice with missing values even with a 'na.rm=TRUE' parameter so a custom solution was coded to handle pesky 'NA' values. In the future, I would probably handle this via an ```addmargins()``` or ```tapply()``` function which hopefully behaves better with 'NA' values.

```{r}
for (i in 1:nrow(reslt)){
   counter=7
   if (is.na(PRERATING[reslt$Opp1[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp2[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp3[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp4[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp5[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp6[i]])) {counter=counter-1}
   if (is.na(PRERATING[reslt$Opp7[i]])) {counter=counter-1}
   reslt$avg.score[i]=round(
      sum(
      PRERATING[reslt$Opp1[i]],
      PRERATING[reslt$Opp2[i]],
      PRERATING[reslt$Opp3[i]],
      PRERATING[reslt$Opp4[i]],
      PRERATING[reslt$Opp5[i]],
      PRERATING[reslt$Opp6[i]],
      PRERATING[reslt$Opp7[i]],
      na.rm=TRUE
      )
   /counter,0)
}
reslt
```
  
### Output CSV. file  

```{r}
write.csv(reslt,"C:\\Users\\LENOVO\\OneDrive\\Learning\\Courses In Progress\\DATA 607\\Projects\\Project 1\\chessresults.csv")
```
  
### So who are the most improved players, really?  
  
We might ask, "who are the most improved players?" or which players have shown the biggest increases between their Pre-rating and Post-rating scores. Let's find out who the top-10 most improved players are by extracting the Post-rating and comparing it to the previously extracted Pre-rating...  
  
```{r}
patrn.POSTRATING="->\\s{0,1}\\d*...."
POSTRATING=unlist(str_extract_all(tour.info,pattern=patrn.POSTRATING))
POSTRATING=POSTRATING[-1]
POSTRATING=str_remove_all(POSTRATING,"->")
POSTRATING=trimws(trimws(POSTRATING,c("left")),c("right"))
POSTRATING=str_remove_all(POSTRATING,"[A-Z][0-9]+")
POSTRATING=as.integer(POSTRATING)
reslt$POSTRATING=POSTRATING
reslt$IMPROVEMENT=reslt$POSTRATING - reslt$PRERATING
reslt3=reslt[order(-reslt$IMPROVEMENT),]
reslt3=reslt3[1:10,c("NAME","PRERATING","POSTRATING","IMPROVEMENT")]

reslt3
```
  
### What are the odds of winning a match?
  
We might also ask, what are the chances of one chess player beating another in a match, and how many rating points would the victor gain in that win? The Elo rating system provides a neat formula to calculate the probability of player a beating player b: ``Pa = 1 / 1 + 10^((Rb - Ra) /400 )`` where ``Ra`` and ``Rb`` are the respective player numerical score ratings. It also provides a formula for determining points gained from such a win, ``K (1 - Pa)`` where ``K`` is a constant. We use K=20 here.  

Let's calculate the odds of and the rating points gained from a win using an R function called ``Chance.Win`` where the first parameter is the player number and the second parameter is the opponent's player number.  
  
```{r}
Chance.Win=function (player,opponent) {
   pw = (POSTRATING[opponent] - POSTRATING[player]) / 400
   prob.win = 1 / (1 + 10^pw)
   
   K=20
   points.gained=round(K * (1 - prob.win))
   
   cat ("The chance of player no.",player, 
        "beating player no.",opponent, "is",
        round(prob.win,2)*100,
        "%,\n","with",points.gained,"point(s) gained in the win.",
        "\n\n")
}

Chance.Win (11,44)
Chance.Win (55,22)
Chance.Win (35,34)
```
  
### How does the outcome of match impact a rating score?  
  
More generally, we can ask: given a win, loss or tie, how do player rating points increase or decrease? We can answer this more general question using an R function called ``Points.Calc`` with an additional parameter ``match.outcome`` which takes the values 'w' for a Win, 'l' for a Loss and 'd' for a Draw.  

```{r}
Points.Calc = function(player,opponent,match.outcome) {
   pw = (POSTRATING[opponent] - POSTRATING[player]) / 400
   prob.win = 1 / (1 + 10^pw)
   
   K=20
   if (match.outcome=="w") {outcome=1
      match.outcome="win"}
   if (match.outcome=="l") {outcome=0
      match.outcome="loss"}
   if (match.outcome=="d") {outcome=0.5
      match.outcome="draw"}
   points.gained=round(K * (outcome - prob.win))
   
   cat ("Player no.",player,"accrues",points.gained,
        "point(s) in a",match.outcome,"against player no.",opponent)
   
}
Points.Calc(2,43,"l")
Points.Calc(54,37,"w")
Points.Calc(37,54,"w")
Points.Calc(27,29,"d")
```
  
  
  
  
  

### But don't take my word for it...

- [This document](https://rpubs.com/MauricioClaudio)

- [Its R Markdown file](https://github.com/MauricioClaudio/CUNY-MPS-DATA-607)
  
- [Its source text data](https://github.com/MauricioClaudio/CUNY-MPS-DATA-607/blob/b9646fc2ef6e9d7e94b1c2f73fa5f295636ec925/tournamentinfo.txt)  
  
- [Its output CSV. file](https://github.com/MauricioClaudio/CUNY-MPS-DATA-607/blob/e82d39970edb7b84ddc40770c3006203b273ac6d/chessresults.csv)  

