---
title: "DATA 607 Assignment 1: Loading Data into a Data Frame"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Where Police Have Killed Americans In 2015
https://fivethirtyeight.com/features/where-police-have-killed-americans-in-2015/
  
#### Overview

Official statistics on police killings in the United States are notoriously unreliable and incomplete. The Guardian, a British daily, compiled data from several different sources and verified them via crowd-sourcing. The dataset is available on GitHub by following the link the article or on my own OneDrive here:  
https://raw.githubusercontent.com/MauricioClaudio/CUNY-MPS-DATA-607/7b0170ebab5907fdc0a9bfe6643e1a746c5a84c9/police_killings.csv

#### Reading in the dataset

Prior to reading in the dataset, the data.table library was loaded. The dataset in CSV format was then read into a data table consisting of 467 observations of 34 variables:

```{r}
library(data.table)
pkdat=data.table(read.csv(
  "C:\\Users\\LENOVO\\OneDrive\\Documents\\Public Code & Data\\police_killings.csv",header=TRUE))
str(pkdat)
```


#### Subsetting the dataset

Once in a data table, eleven unnecessary or superfluous columns were eliminated. The following command keeps all rows for columns with names other than those within the c() vector. Said another way, columns within the c() vector were eliminated. The resulting subset dataset consists of 467 observations of  23 variables.

```{r}
pkdat.subset=pkdat[,-c(
  "name",
  "month",
  "day",
  "year",
  "streetaddress",
  "city",
  "state",
  "latitude",
  "longitude",
  "lawenforcementagency",
  "comp_income")]
str(pkdat.subset)
```


#### Renaming columns in the dataset

The third and last step was to rename columns with ambiguous or unclear names. This was accomplished by creating a vector with the original column names to be replaced and a vector with the replacement column names, and using the two vectors as the second and third arguments of the setnames function


```{r}
library(data.table)
name.old=c(
  "state_fp",
  "county_fp",
  "cause",
  "p_income",
  "h_income",
  "urate",
  "tract_ce",
  "geo_id",
  "namelsad",
  "county_bucket",
  "nat_bucket",
  "pov",
  "college"
  )
name.new=c(
  "stateFIPScode",
  "countyFIPScode",
  "cause_of_death",
  "mean.income.personal",
  "mean.income.household",
  "unemployment%",
  "TractIDcode",
  "CombinedTractIDcode",
  "TractDescription",
  "HouseholdIncomeQuintileCounty",
  "HouseholdIncomeQuintileNationally",
  "TractPovertyRate",
  "Share25+popWithBAorHigher"
  )
setnames(pkdat.subset,name.old,name.new)
str(pkdat.subset)
```


#### Findings & Follow-up

The data for the year 2015, augmented with US Census data, shows that deaths at the hands of the police take place primarily, though not exclusively, in lower-income, economically disadvantaged, African-American neighborhoods. The dataset retains the earlier 467 observations of 23 variables.

The data suggests that drivers of police brutality correlate with income and education. To explore possible alleviation interventions comprising drivers and inhibitors, I would look for locations where no killings took place yet sharing similar economic and educational indicators to those where killings took place. Then, I would try to figure out the factors in those localities that acted as inhibitors.

Likewise, I would compare these figures with other countries. I recall reading that in Brazil's Rio de Janeiro, a city of less than seven million inhabitants, more people die at the hands of police in a year than in the whole of the United States (population 333 million)! Police brutality in the United States is only the tip of a worldwide iceberg.


